
\paragraph{Exposure}
The performativity of databasing can be understood in terms of what Nancy calls exposure \parencite{Nan91:The}. 

% this is not a great definition of exposure. Try to define it in your own words. you already have the quote to define it in Nancy's words.

Exposure is the appearance of a limit and the finitude of a singularity. With this limit instantiated in the (public) moment of the performative act, is how communication emerges as that which is in common among singularities. That is to say, because it is itself nothing (``neither a ground, nor an essence, nor a substance'' \parencite[31]{Nan91:The}) Nancy considers finitude to just appear in the form of communication: ``it presents itself, it exposes itself, and thus it exists as communication'' \parencite[31]{Nan91:The}. 

% Alright, as soon as nancy wnters we have some trouble. I wonder if this perception changes after your rewriting of this section on unworking earlier, but I need to move on at this point. can you revise this section. you are making claims, but many of these claims seem abstract and out of context. Why does it matter if "the database" is an agent of community? how does this change anything?

His emphasis on communication as exposure marks a crucial distinction on the concept of community. For Nancy, as I have described above \see{inoperativity}, community cannot come from an instance of work: it emerges as an instance of the communicative action which is ``the unworking of a work that is social, economic, technical, and institutional'' \parencite[31]{Nan91:The}. In other words, the performativity of this communicative act, the publicness and dramatic qualities with which it unfolds, result in exposure. This is why, in the case of databasing, what is exposed at the limit of its performance is the finitude of the database itself. Through this exposition of the limit is how the singularity of the database can be communicated; or, better, through this exposure is how communication expands our concept of community towards one that includes the database as an agent of community itself.

\paragraph{Anarchic Touch}
% try re-writing. To many does not, is not, comes not. Try something more affirmative or restructure so that it doesn't read as a list of what is nots
% Also, make sure you're being clear. you have a tendency to get swallowed in complexity when you are explaining things.

As Nancy writes, ``a singular being does not emerge or rise up against the background of a chaotic, undifferentiated identity of beings'' \cite[28]{Nan91:The}. This is to say that, like Butler's gendered self, there is no substance within singularity. 
% this wor dkeeps showing u, but it is either not well defined or not defined at all.
The appearance of a database as finitude comes not from an originary \textit{archē} which would impose its archontic power as is the case of archives. 
% but rather...
The limit of the database is not instantiated out of the one, a ``unitary assumption'' \cite[28]{Nan91:The}, or the wholeness of a single one. This finitude does not come from intentionality or any essentialist notions: it simply appears, for Nancy, in the form of touch: ``\dots as finitude itself: at the end (or at the beginning), with the contact of the \textit{skin} (or the heart) of another singular being, at the confines of the same singularity that is, as such, always other, always shared, always exposed\dots'' \im \cite[28]{Nan91:The}. The temporality of this touch, like that of gender, is anarchic.

\paragraph{Communities of Skin}
% this section is a bit more clear. It reminds me of the idea of interface as skin. perhaps you might make that connection earlier? otherwise skin remains always abstract.

The limit of the database, as performative, spectral skin, allows for a community to emerge between the human and the nonhuman. This means that the agency locus of the database needs to be placed precisely on its skin, because it is what becomes public of itself. In other words, given that this skin is available to the perception of others, it becomes touchable, it reaches our own limit as databasers. By exposing our own limits to ourselves and to each other, the database changes our definition often delimits the extent of our own singularity. 
% it it it
However, this does not mean that it stands in the way of our performativity, or worse, that it precludes or determines ourselves. If this were true, we would be once again subject to technical determinism, essence fabrication, etc., and falling out of considering any possibility of community between anything that is not human 
% i talked about this! remove
 As Nancy rightfully claims, `` it is not obvious that the community of singularities is limited to `man''' \cite[28]{Nan91:The}. Thus, the fact that the skin of the database changes our own skin simply means that we are already in communication with it, that is, in community, and also in a state of resonance with it. This is the function of the skin of the database: like the skin of a drum, or the skin of a loudspeaker, the skin of the database resonates with our own skin, engaging the resonant body with a resonant spectrality. 

 % bad sentence structure
This is the community of resonance, the community of the resonant network, which has no purpose, no intentionality behind, no essence; only appearance and motility, performance and repetition: an activity that is 
% this keeps somewhat hard to understand. Skin is now very clear after this paragraph, but unworking is still lost in abstraction to me
the unworking with which community exposes itself. 

\paragraph{Hybrid Pluralities}
Database models tend to reside next to each other, either within a single database system or within an interconnected networked system. Databasers have access to the many features that each model offers, focusing on those features that are suitable for their needs.  The skin of the database is as fluid as the constitution of gender, and if this is true, then the fluidity of databasing itself comes to represent the constitution of gender through the performativity of databasers. By resonanting in such performativity, databasers approach (but do not reach) the limit of the database. This approach to the skin of the database exposes simultaneously 
% simultaneously, what and what?
the skin of the databaser to the database. 
%
% too abstract ?
What this exposure amounts to is not, however, an opposition of forces. It results in the fragmented state of community that resides in the different degrees of this exposure. In other words, this exposure is of a hybrid plurality that resonates at the limit. Engaging with the touch of the spectral database means reconfiguring, resounding, and remembering our own sense of touch, just as well as our own sense of self.