Derrida's broader notion of `writing' relates to what he calls an `originary' trace. As I described earlier, the inscription of the trace contains from the start its own erasure. This means that, in a reciprocal motion, the origin of the trace is at once originary and non-originary: the origin of a disappearance and the disappearance of the origin. This apparent contradiction can be approached by the actual process of writing something on a page: where or when does the writing begin, when the pen touches the paper or when the paper lends itself for writing? The same question applies to the crossfade between the `I's in the case of Alvin Lucier's work \see{network}, where or when does the `I' begin `sitting' in the loop? Further, no matter how many metaphors we may give to this paradox, it simply falls out of empirical quests. Therefore, this paradox of the trace is understood by the concept of \textit{différance} that we outlined above. The repercussions of this word that ``is not a word and it not a concept'' have now a long history that does not interest us fully, therefore we can will simply jump thirty years ahead in time, after computers entered the scene, and after the Internet became popularized by the World Wide Web.

\textcite{Der95:Arc} exemplified the intricacies that come out of the process of archivization, delineating an economy of archives into he calls the \textit{archontic} principle.\footnote{His example was the Sigmund Freud Museum, which is located in the same house where Freud lived.\url{https://www.freud-museum.at/en/}} A brief etymological impasse, coming from Derrida's text, might be pertinent here. `Economy' is related to the greek word \textit{oikos} (house), and \textit{nomos} (man-made law). In ancient Greece, the official documents of the law were kept under guard in the house of the magistrates (\textit{arkheion}), in a form of house arrest. Funes, as I have described earlier \see{funeslude}, was in a similar form of house arrest (arrested from the outside world). The archive condenses this economy, however, within the confines of a public place: by an institutional passage from the private (house) to the public (archive, museum). The word `archive' (\textit{arkheion}) comes from the greek \textit{archē} which relates, on the one hand, to the originary, as well as to the ruling. Thus, the archontic principle is a type of authority that the archive exerts, which can be understood as the law of the house, or the law that is before anything else. Hence, its categorization as principle, which is also related to the origin (e.g., the latin root \textit{principium} which refers to the beginning) and to the figure of the ruler (principal or prince). As hypomnesis, archives represent for Derrida another instance of the movement of technology. This movement consists in ``a transformation of the techniques of archivization, of printing, of inscription\dots'' \parencite[16]{Der95:Arc}. These `archival machines' which developed in the thirty years since he first considered Freud's \textit{Note on the Mystic Pad} had reached Derrida's email inbox: his comments on how email technology would have (and will) affect psychoanalysis attest to this fact. Further, they pose a question that he leaves nonetheless unanswered, which is why I consider his discussion of the archive relevant to our discussion of database aesthetics. Derrida finds in Freud's metaphor for the psychic apparatus a point of departure to the question of how psychoanalysis changed by the presence of a technological device. In this case, the device was a toy called the Mystic Pad, a children's writing board made of wax and a thin layer on top: upon impression it leaves a trace; when one lifts the layer, the trace erases. The structure of the psyche for Freud could thus be understood by using both hands simultaneously, one pressing (writing), while the other was lifting (erasing). Writing in 1995, Derrida posed the following question:

\begin{quote}
	Is the psychic apparatus better represented or is it affected differently by all the technical mechanisms for archivization and for reproduction, for prostheses of so-called live memory, for simulacrums of living things which already are, and will increasingly be, more refined, complicated, powerful than the `mystic pad' (microcomputing, electronization, computerization, etc.)? \parencite[16]{Der95:Arc} 
\end{quote}

% I will begin to approach this question in two ways, but first I would like to narrow down the scope of its premises. On the one hand, the `psychic apparatus' only interests us in relation to what I have mentioned of memory, and more precisely of a certain quality of the unconscious (death drive). On the other, the techno-science or technical mechanism that is in question here is that of the database. Thus, I will proceed to break the question appart, and then turn it upside down, so as to bring it into the sphere of database aesthetics. 

The structure of this question points to two possible answers. Either technology is a representation of the psyche, or the understanding of the psyche is changed by technology. On one hand, the question refers to how distant we are in the movement with which we can arrive at the object (psychic apparatus). This is a temporal movement which relates to deferral: how much longer until we get our object? The answer, as we have seen, is simply never. In order for the represented to be fully represented it must become present. On the other hand, the `technical device' has the potential to affect the object: it differs from and might reformulate thus the psychic apparatus. To a certain extent, this suggests that these `prosthesis' of `live' memory, or `simulacrums of living things' represent some form of (disembodied) nonhuman (life) against which the question is asked. We can only guess. In any case, Derrida claims it is a question of progress or evolution, in which ``neither of these hypothesis can be reduced to the other'' \parencite[16]{Der95:Arc}. It appears, therefore, that this quest reaches a dead end. Derrida does not continue on this quest, and instead takes the text in directions that are not pertinent for our purposes here. Therefore we can stress on this question for a while. 

The structure of the memory (``the essence of the psyche'') in opposition to the archive can be understood with the opposition of \textit{anamnesis} (the act of recalling, remembering) and \textit{hypomnesis} (the technical storage device); the former internal, the latter external. We can advance now that, like the archive, the database is hypomnesis and external (to the psyche), just as well as it exerts the principles of the archontic. As I have described earlier \see{chapter:part-2}, a database comprises the partition within computer memory where data is stored. However, in order to store data, one has to assign data types and structures, thus providing with the necessary structural frame that indicates how to access the data. One of the key concepts of the `archontic principle' is `consignation,' which has at least three meanings: assigning residence, entrusting something in reserve or deposit, and what Derrida calls ``gathering together signs'' \parencite[10]{Der95:Arc}. All of these can be represented with memory management, that is, by `declaring' and `initializing' variables, as well as by making `unions' or more complex data structures. Consignation in archives has, for Derrida, a presupposing aim: to ``coordinate a single corpus'' \parencite[10]{Der95:Arc}. This is exactly the case of a database, which extends this coordination with the possibility of networks. The various database models that I have shown in previous sections point to the different ways with which to coordinate the economy of database systems. In order to flip Derrida's question upside down, we can ask ourselves what to do with all these technical mechanisms that microcomputing enables? How do we make something out of them? And further, what is in them that we can call aesthetic? Instead of thinking how memory is represented in the database, we can ask how we can represent the database within memory. Further, instead of asking how memory is affected by the database, we can ask ourselves how can memory affect databases. The theoretical and aesthetical aspects of this reversal can be seen as follows. On one hand, this reversal points to a reconceptualization of databases within the ``evolution of archival techno-science'' \parencite[16]{Der95:Arc}. On the other, this reversal calls for a restructuring or a technical reconstruction of the database in relation to memory. 


In his architecture, Von Neumann proposed that the storage unit of the computer would allow for data to be written and erased in different locations and times. He was following Turing's conceptualization of the \textit{a-machine} ---i.e., the \textit{Turing machine}---, which was a mathematical model for computation, that can be represented by a symbol scanner and an infinite tape, where the scanner gets, sets, or unsets a symbol on the tape, and the tape moves to the next slot accordingly. These setting and unsetting movements represent inscription and erasure, to the point that, as Kittler notes: ``the two most important directing signals which link the central processing unit of the computer to external memory are being called \texttt{READ} and \texttt{WRITE}'' \parencite[131]{Ern13:Dig}. An important distinction needs to be made here. While I am arguing for the similarities that exist between human memory and databases, Ernst instead proposes that databases or ``digital an-archives'' replace human memory. As media tend to converge toward digital media, `reading’ gives way to mathematical processes that interpret data: ``signal processing replaces \textit{pure} reading'' \im \parencite[130]{Ern13:Dig}. This statement resonates with the Kittlerian, disembodied worldview \see{convergence}, could be extended to saying that signal processing replaces \textit{listening}. 


Convergence, understood in this way, points to a world view where data structures and algorithms replace our own memory: the ultimate convergence is the absorption of human memory (the psychical apparatus) in the database. In this sense, the database would remove our embodiment from our bodies completely. There is a fine line between this convergence, and what Derrida mentions about consignation in relation to archives. In both terms, there is an aim that is presupposed: everything goes in the same place, archive or database, and from this technological place we find a hint of our own destruction as humans. That is to say, the Kittlerian residue of humanity understood as an expression of the pull towards the inorganic, the death drive, the archive `fever.' But, if the database reads itself, it also writes itself, and, in the case of music, if it listens to itself, it also sounds itself, and we, as humans, simply are part of this loop and of its resonance. This moment of resonance enables both the nonhuman and the human to coexist. Of course, the paradox of a destructive force that is only possible through its destruction might be no welcomed in our houses, in our databases, and in ourselves. What this resistance indicates is the presence of an absence, a sound without a body, and the spectrality of an uninvited guest: and it is the sound of this force that makes an impression. Derrida calls these ``lovely impressions'' or ``memories of death'' \parencite[14]{Der95:Arc}, relating the death drive to the anarchic and an-archontic force that leaves behind no trace because it ``always operates in silence'' \parencite[14]{Der95:Arc}. Under these terms, could we not reconfigure this fever in relation to databases as a force of inoperativity, one that unworks in resonance?


